ML Estimation of Matrix--$t$ Distributions with Expectation Maximization

### 2017-01-21 paper.tex
Finished checking the E-step and corrected some minor typos.

\begin{itemize}
\item At the $k$-th iteration, compute the conditional expectations with respect to $Z$
\begin{equation}
Q(\theta,\theta_k):=\sum_{i\in I}\mathbb{E}_{\theta_k}\left[\log p_{\theta}(x_i,Z)|x_i\right].
\end{equation}

@@

We can split the integral in Eq.~\eqref{eq:expdef} by the terms of $\log f(T_{j},S;\theta)$. Eliminating a global factor $\frac{1}{2}$ and terms which do not depend on the parameters in $\theta$ we have
\begin{equation}
\begin{split}\log f(T_{j},S;\theta)\sim -\nu\log|S|-\textrm{tr}\left[S^{-1}(\Sigma+(T_j-M)\Omega^{-1}(T_j-M)^{\textrm{t}})\right]\\
-\nu p\log2-p\log |\Omega|+(\nu+p-1)\log |\Sigma|-2\log\Gamma_{p}\left(\frac{\nu+p-1}{2}\right).
\end{split}\label{eq:logf}
\end{equation}
From Eq.~\eqref{eq:logf}, we can see the only integrals which need to be computed are
\begin{equation}
\int_{\mathbf{S}_+^p}f(T_j,S;\theta_k)\log |S|\,dS\label{eq:firstlogpart}
\end{equation}
and
\begin{equation}
\int_{\mathbf{S}_+^p}f(T_j,S;\theta_k) S^{-1}\,dS.\label{eq:2ndlogpart}
\end{equation}
Note that taking the derivative with respect to $\alpha$ in Eq.~\eqref{eq:intmultgammadef} we have
\begin{equation}
\int_{\mathbf{S}_+^p}e^{-\textrm{tr}\left[SA\right]}|S|^{\alpha}\log |S|\,dS=\Gamma_{p}\left(\alpha+\frac{1+p}{2}\right)|A|^{\alpha+\frac{1+p}{2}}\left\{\log |A|+\gamma_{p}\left(\alpha+\frac{1+p}{2}\right)\right\}\label{eq:intderivative1}
\end{equation}
and hence the integral \eqref{eq:firstlogpart} follows this result directly. To compute the integral \eqref{eq:2ndlogpart} we take the matrix derivative of Eq.~\eqref{eq:intmultgammadef} with respect to $A$ and obtain
\begin{equation}
\int_{\mathbf{S}_+^p}e^{-\textrm{tr}\left[SA\right]}|S|^{\alpha}S\,dS=-\left(\alpha+\frac{1+p}{2}\right)\Gamma_{p}\left(\alpha+\frac{1+p}{2}\right)|A|^{\alpha+\frac{p-1}{2}}A^{-1},\label{eq:intderivative2}
\end{equation}
from where again, we can obtain the result by direct substitution.

Combining results of Eqs.~\eqref{eq:intderivative1} and \eqref{eq:intderivative2} with adequate rearrangements we obtain the closed form of $Q_j(\theta,\theta_k)$ while discarding the irrelevant constant terms
\begin{equation}
\begin{split}Q_j(\theta,\theta_k)\sim \nu\psi_{p}\left(\frac{m+\nu_k+p-1}{2}\right)-2\log\Gamma_{p}\left(\frac{\nu+p-1}{2}\right)+p\log|\Omega|^{-1}\\
-\nu\log|(T_j-M_k)\Omega_k^{-1}(T_j-M_k)^{\textrm{t}}+\Sigma_k|+(\nu+p-1)\log |\Sigma|\\
-(m+\nu_k+p-1)\textrm{tr}\left[\left((T_j-M_k)\Omega_k^{-1}(T_j-M_k)^{\textrm{t}}+\Sigma_k\right)^{-1}\right.\\
\left.\left((T_j-M)\Omega^{-1}(T_j-M)^{\textrm{t}}+\Sigma\right)\right].
\end{split}
\label{eq:fsimp}
\end{equation}
The maximization formulae is obtained setting the derivatives of this equation to zero.

>>>>

\begin{itemize}
\item At the $k$-th iteration, compute the conditional expectations with respect to $Z$
\begin{equation}
Q(\theta,\theta_k)=\sum_{i\in I}Q_i(\theta,\theta_k),\qquad Q_i(\theta,\theta_k)=\mathbb{E}_{\theta_k}\left[\log p_{\theta}(x_i,Z)|x_i\right].
\end{equation}

@@

We can split the integral in Eq.~\eqref{eq:expdef} by the terms of $\log f(T_{j},S;\theta)$. Eliminating a global factor $\frac{1}{2}$ and terms which do not depend on the parameters in $\theta$ we have
\begin{equation}
\begin{split}\log f(T_{j},S;\theta)\sim -\nu\log|S|-\textrm{tr}\left[S^{-1}\left\{\Sigma+(T_j-M)\Omega^{-1}(T_j-M)^{\textrm{t}}\right\}\right]\\
-\nu p\log2-p\log |\Omega|+(\nu+p-1)\log |\Sigma|-2\log\Gamma_{p}\left(\frac{\nu+p-1}{2}\right).
\end{split}\label{eq:logf}
\end{equation}
The integrals to be computed are
\begin{equation}
\int_{\mathbf{S}_+^p}f(T_j,S;\theta_k)\log |S|\,dS, \qquad 
\int_{\mathbf{S}_+^p}f(T_j,S;\theta_k) S^{-1}\,dS
\end{equation}
and they can be done by using the equalities derived by taking the derivative of Eq.~\eqref{eq:intmultgammadef} with respect to $\alpha$ and matrix $A$, respectively,
\begin{equation}
\int_{\mathbf{S}_+^p}e^{-\textrm{tr}\left[SA\right]}|S|^{\alpha}\log |S|\,dS=\Gamma_{p}\left(\alpha+\frac{1+p}{2}\right)|A|^{-\alpha-\frac{1+p}{2}}\left\{\gamma_{p}\left(\alpha+\frac{1+p}{2}\right)-\log |A|\right\},\label{eq:intderivative1}
\end{equation}
\begin{equation}
\int_{\mathbf{S}_+^p}e^{-\textrm{tr}\left[SA\right]}|S|^{\alpha}S\,dS=\left(\alpha+\frac{1+p}{2}\right)\Gamma_{p}\left(\alpha+\frac{1+p}{2}\right)|A|^{-\alpha-\frac{1+p}{2}}A^{-1}.\label{eq:intderivative2}
\end{equation}

Utilizing the equalities \eqref{eq:intderivative1} and \eqref{eq:intderivative2} with adequate rearrangements we obtain the closed form expression of $Q_j(\theta,\theta_k)$ while ignoring the irrelevant constant terms
\begin{equation}
\begin{split}Q_j(\theta,\theta_k) \sim \nu\psi_{p}\left(\frac{m+\nu_k+p-1}{2}\right)-2\log\Gamma_{p}\left(\frac{\nu+p-1}{2}\right)-p\log|\Omega|\\
-\nu\log|\Sigma_k+(T_j-M_k)\Omega_k^{-1}(T_j-M_k)^{\textrm{t}}|+(\nu+p-1)\log |\Sigma|\\
-(m+\nu_k+p-1)\textrm{tr}\left[\left\{\Sigma_k+(T_j-M_k)\Omega_k^{-1}(T_j-M_k)^{\textrm{t}}\right\}^{-1}\right.\\
\left.\left\{(T_j-M)\Omega^{-1}(T_j-M)^{\textrm{t}}+\Sigma\right\}\right].
\end{split}
\label{eq:fsimp}
\end{equation}
The maximization formulae follows by setting the gradient of the parameters to zero.
